---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# DeconvoBuddies

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/lahuuki/DeconvoBuddies/branch/main/graph/badge.svg)](https://codecov.io/gh/lahuuki/DeconvoBuddies?branch=main)
[![R-CMD-check-bioc](https://github.com/lahuuki/DeconvoBuddies/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/lahuuki/DeconvoBuddies/actions)
<!-- badges: end -->

The goal of `DeconvoBuddies` is to provide helper functions for the deconvolution process

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `DeconvoBuddies` using from [Bioconductor](http://bioconductor.org/) the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("DeconvoBuddies")
```

And the development version from [GitHub](https://github.com/) with:

```{r 'install_dev', eval = FALSE}
BiocManager::install("lahuuki/DeconvoBuddies")
```
## Example

```{r 'load packages', echo = FALSE}
library("DeconvoBuddies")
library("dplyr")
library("ggplot2")
```


Get mean ratios for each gene x cell type
```{r example, eval = requireNamespace('DeconvoBuddies')}
ratios <- get_mean_ratio2(sce.test)
fc <- findMarkers_1vAll(sce.test)

(marker_stats <- left_join(ratios, fc, by = c("gene", "cellType.target")))
```

### Extablish Color Scheme
```{r establish_colors}
cell_types <- levels(sce.test$cellType)
cell_colors <- create_cell_colors(cell_types = cell_types, pallet = "classic", split = "\\.", preview = TRUE)
```


### Plot Expression of Marker Genes
```{r plot_marker_expression, echo = FALSE}
plot_marker_express(sce.test, marker_stats, "Astro",n_genes = 5,rank_col = "rank_ratio", anno_col = "anno_ratio") +
    scale_color_manual(values = cell_colors)
```

### Create Composition Bar Plot
```{r}
pd <- SummarizedExperiment::colData(rse_bulk_test) %>%
     as.data.frame()

 est_prop_long <- est_prop %>%
     tibble::rownames_to_column("RNum") %>%
     tidyr::pivot_longer(!RNum, names_to = "cell_type", values_to = "prop") %>%
     dplyr::left_join(pd %>% dplyr::select(RNum, Dx)) %>%
     dplyr::mutate(a = "a")

 plot_composition_bar(est_prop_long)
 plot_composition_bar(est_prop_long, x_col = "Dx")
 plot_composition_bar(est_prop_long, x_col = "Dx", min_prop_text = 0.1)
```



## Citation

Below is the citation output from using `citation('DeconvoBuddies')` in R. Please
run this yourself to check for any updates on how to cite __DeconvoBuddies__.

```{r 'citation', eval = requireNamespace('DeconvoBuddies')}
print(citation('DeconvoBuddies'), bibtex = TRUE)
```

Please note that the `DeconvoBuddies` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing this package.

## Code of Conduct

Please note that the `DeconvoBuddies` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://.github.io/DeconvoBuddies) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.


