---
title: "Deconvolution_Benchmark_DLPFC"
author: 
  - name: Louise Huuki-Myers
    affiliation:
    - &libd Lieber Institute for Brain Development, Johns Hopkins Medical Campus
    email: lahuuki@gmail.com
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('DeconvoBuddies')`"
vignette: >
  %\VignetteIndexEntry{Deconvolution_Benchmark_DLPFC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```


```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## Track time spent on making the vignette
startTime <- Sys.time()

## Bib setup
library("RefManageR")

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    knitr = citation("knitr")[1],
    RefManageR = citation("RefManageR")[1],
    rmarkdown = citation("rmarkdown")[1],
    sessioninfo = citation("sessioninfo")[1],
    testthat = citation("testthat")[1],
    DeconvoBuddies = citation("DeconvoBuddies")[1]
)
```

#Introduction

## What is Deconvolution?

**Inferring the composition of different cell types in a bulk RNA-seq data**

`Bisque` is an R package for preforming reference based deconvolution [github.com/cozygene/bisque](https://github.com/cozygene/bisque).

## Goals of this Vignette 

We will be demonstrating how to use `DeconvoBuddies` tools when applying 
deconvolution with the `Bisque` package.

1. Install and load required packages 
2. Download DLPFC RNA-seq data, and reference snRNA-seq data
3. Find marker genes with `DeconvoBuddies` tools
4. Run deconvolution with `BisqueRNA`
5. Explore deconvolution output and create compostion plots with `DeconvoBuddies` tools
6. Check proportion against RNAScope estimated proportions

# Basics

# 1. Install and load required packages

`R` is an open-source statistical environment which can be easily modified to enhance its functionality via packages. `r Biocpkg("DeconvoBuddies")` is a `R` package available via the [Bioconductor](http://bioconductor.org) repository for packages. `R` can be installed on any operating system from [CRAN](https://cran.r-project.org/) after which you can install `r Biocpkg("DeconvoBuddies")` by using the following commands in your `R` session:

## Install `DeconvoBuddies`

```{r "install", eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
      install.packages("BiocManager")
  }

BiocManager::install("DeconvoBuddies")

## Check that you have a valid Bioconductor installation
BiocManager::valid()
```

## Load Other Packages

```{r "load_packages", message=FALSE, warning=FALSE}
## install Bisque from cran
# install.packages("BisqueRNA")

library("spatialLIBD")
library("DeconvoBuddies")
library("SummarizedExperiment")
library("SingleCellExperiment")
library("BisqueRNA")
library("dplyr")
library("tidyr")
library("tibble")
```

# 2. Download DLPFC RNA-seq data, and reference snRNA-seq data. 

## Bulk RNA-seq data

Access the 110 sample Human DLPFC bulk RNA-seq dataset for LIBD. These samples  
are from 19 tissue blocks, and 10 neurotypical adult donors. Samples were sequenced
with two different `library_types` (polyA and RiboZeroGold), and three different 
`RNA_extraction` (Cyto, Total, Nuc), post quality control n=110 samples. 

```{r "load rse_gene"}
## use fetch deconvon data to load rse_gene
rse_gene <- fetch_deconvo_data("rse_gene")
rse_gene

## bulk RNA seq samples were sequenced with different library types, and RNA extractions
table(rse_gene$library_type, rse_gene$library_prep)
```

## Refernce snRNA-seq data

This data is paired with a single nucleus RNA-seq data set from `spatialLIBD`. 
This dataset can be accessed with `spatialLIBD::fetch_data()`.

```{r "load snRNA-seq"}
## Use spatialLIBD to fetch the snRNA-seq dataset
sce_path_zip <- spatialLIBD::fetch_data("spatialDLPFC_snRNAseq")

# sce_path <- unzip(sce_path_zip, exdir = tempdir())
# 
# sce <- HDF5Array::loadHDF5SummarizedExperiment(
#     file.path(tempdir(), "sce_DLPFC_annotated")
# )
```


# 3. Find marker genes with `DeconvoBuddies` tools
# 4. Run deconvolution with `BisqueRNA`
# 5. Explore deconvolution output and create compostion plots with `DeconvoBuddies` tools
# 6. Check proportion against RNAScope estimated proportions


# Reproducibility

The `r Biocpkg("DeconvoBuddies")` package `r Citep(bib[["DeconvoBuddies"]])` was made possible thanks to:

* R `r Citep(bib[["R"]])`
* `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
* `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])`
* `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`
* `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])`
* `r CRANpkg("sessioninfo")` `r Citep(bib[["sessioninfo"]])`
* `r CRANpkg("testthat")` `r Citep(bib[["testthat"]])`

This package was developed using `r BiocStyle::Biocpkg("biocthis")`.


Code for creating the vignette

```{r createVignette, eval=FALSE}
## Create the vignette
library("rmarkdown")
system.time(render("Deconvolution_Benchmark_DLPFC.Rmd", "BiocStyle::html_document"))

## Extract the R code
library("knitr")
knit("Deconvolution_Benchmark_DLPFC.Rmd", tangle = TRUE)
```

Date the vignette was generated.

```{r reproduce1, echo=FALSE}
## Date the vignette was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r reproduce2, echo=FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```



# Bibliography

This vignette was generated using `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])` and `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])` running behind the scenes.

Citations made with `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`.

```{r vignetteBiblio, results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```
